Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-11-10 ms.tex
Rephrasing some things in discussion of Harrington separation of variables

Approximate analytic solutions for resonant scattering exist in certain limits. \citet{1973MNRAS.162...43H} showed that when most of the radiation is out on the damping wings, the transfer equation reduces to the Poisson equation. However, their solution contains an incorrect treatment of the boundary condition. To our knowledge, the errors introduced by this treatment have never been quantified. They attempt a separation of variables $J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant is shown to satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi \Delta,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, $\phi$ is the line profile, and $\Delta$ is the Doppler width. The key point is that the line profile depends on one of the coordinates, frequency, and this causes the eigenvalues of the separation constant to depend on frequency. Thus, the separation ``constant'' is not constant, and the function does not satisfy the Poisson equation since the frequency derivatives will act on the separation ``constant" giving extra terms. In the limit of large optical depth $B$, they approximate the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$. However, this solution gives zero mean intensity at the surface. Their Equation 34 subsequently allowed $\lambda$ to have a small deviation from the above expression, which was explicitly frequency dependent. This allowed a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Our treatment, using the correct boundary condition, quantifies the errors in their ansatz.

>>>>

Approximate analytic solutions for resonant scattering exist in certain limits. \citet{1973MNRAS.162...43H} showed that when most of the radiation is out on the damping wings, the transfer equation reduces to the Poisson equation. However, their solution contains an incorrect treatment of the boundary condition. To our knowledge, the errors introduced by this treatment have never been quantified. They attempt a separation of variables $J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant is shown to satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi \Delta,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, $\phi$ is the line profile, and $\Delta$ is the Doppler width. The key point is that the line profile depends on one of the coordinates: frequency. This causes the eigenvalues of the separation constant to be frequency-dependent. Thus, the separation ``constant'' is not constant, and the function does not satisfy the Poisson equation since the frequency derivatives will act on the separation ``constant", giving extra terms. In the limit of large optical depth $B$, they approximate the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$. However, this gives zero mean intensity at the surface. Their Equation 34 subsequently allows $\lambda$ to have a small deviation from the above expression, which is explicitly frequency dependent. This allows a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Our treatment, using the correct boundary condition, quantifies the errors in their ansatz.
